---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The dataset we are analyzing is the library checkout data from the Seattle Public Library (SPL) from 2017 to 2023. The dataset contains information about the items that have been checked out from the library, including the checkout date, checkout type (e.g., OverDrive, Hoopla), and material type (e.g., book, audiobook). Our goal is to analyze the data to identify any trends in the type of materials being checked out and the checkout types that are most popular. We are primarily interested in analyzing the trends in the checkout types (OverDrive, Hoopla) and the material types (books, audiobooks) over time. We want to identify any changes in the popularity of these checkout types or materials over the years. We are also interested in analyzing the total number of checkouts for each checkout type and material type to determine which items are the most popular among library users.

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

- **What is the average number of checkouts for each item?**
- **What is the month or year with the most/least checkouts for a book that you're interested in?**
- **What is the month or year with the most/least checkouts for ebooks?**
- **How has the number of print book checkouts changed over time?**

Feel free to calculate and report values that you find relevant.

### The Dataset

- **Who collected/published the data?**
  The Seattle Public Library collected and published the dataset used in this project
- **What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?**<br>
  The paramaeters are the following:<br>
    UsageClass: This determines if the data entry for a checkout was a physical or digital item<br>
    CheckoutType: This is the system on which the item was checked out on<br>
    MaterialType: This is the type of item being checked out (ex. book, ebook, magizine, etc)<br>
    CheckoutYear: The year the item was checked out<br>
    CheckoutMonth: The month the item was checked out<br>
    Checkouts: The number of checkouts that specfic item entry had<br>
    Title: Title of the item being checked out<br>
    Creator: Author of the item<br>
    Subjects: Genre of the item<br>
    Publisher: Publisher of the item<br>
    PublicationYear: Year the item was published<br>
- **How was the data collected or generated?**  
  
- **Why was the data collected?**  
  To keep track of a monthly count of Seattle Public Library checkouts by title for physical and electronic     items
- **What, if any, ethical questions do you need to consider when working with this data?**  

- **What are possible limitations or problems with this data?(at least 200 words)**
  There are many limiations/problems with this dataset. 

### First Trends Over Time Chart

``````{r, echo = FALSE, warning = FALSE, code = readLines("Line_Chart.R")}
```

### Second Trends Over Time Chart

``````{r, echo = FALSE, code = readLines("Line_Chart2.R")}
```

### Your Choice

``````{r, echo = FALSE, code = readLines("Bar_Chart.R")}
```


